---
title: "DiffBind & ChIPseeker"
output: html_document
date: "2024-07-12"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

### load library
```{r, warning=FALSE, message = FALSE}

library(DiffBind)
library(tidyverse)
library(stringr)
library(tidyr)
library(DESeq2)
library(ChIPseeker)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)

```

### DiffBind
Output csv files with columns of Chr, Start, End, Conc, Conc_CTd180, Conc_CTd0, Fold, p-value, FDR.
```{r, warning=FALSE, message=FALSE}

#read in sample sheet
sampleTable <- read.csv(file = "data_for_import/CUT&Tag/CUT&TagSampleSheet.csv")
#change col name to fit the diffbind design
colnames(sampleTable)[colnames(sampleTable) == "CellSet"] <- "Factor"

histone_list=c("K4me3", "K27ac", "K27me3")

mark <- "K4me3"

for (mark in histone_list){
  
  #subset sample sheet for each histone mark
  sampleTable_more <- sampleTable[sampleTable$Histone == mark, ]

  ######
  ###(To run DiffBind from scratch, please download the bam files and peak files from GEO database to "data_for_import/CUT&Tag/bam/" and "data_for_import/CUT&Tag/peak" folder respectively, please also make sure the file names match according to the sample sheet above; Alternatively, you can load in RData with all the peak information stored in it and continue do analysis with it.)###
  ######
  
  ##sampleSheet
  #histone <- dba(sampleSheet=sampleTable_more)

  ##blacklist
  #histone <- dba.blacklist(histone, blacklist=DBA_BLACKLIST_GRCH38, greylist=FALSE)

  ##count
  #histone <- dba.count(histone)

  ##normalize
  #histone <- dba.normalize(histone)
  
  #save RData file
  #dba.save(histone, paste(, mark, "Peaks_default", sep = "_"), dir = "output/CUT&Tag/DiffBind")
  
  #load RData file
  histone <- dba.load(file = paste(mark, "Peaks_default", sep = "_"), dir = "output/CUT&Tag/DiffBind")
  
  #contrast and analyze
  histone <- dba.contrast(histone, design="~ Factor + Condition", minMembers=2, reorderMeta = list(Condition="CTd0"))
  dba.contrast(histone, bGetCoefficients=TRUE)
  histone <- dba.contrast(histone, design="~ Factor + Condition", contrast = c("Condition", "CTd120", "CTd0"))
  histone <- dba.analyze(histone)
  histone <- dba.contrast(histone, design="~ Factor + Condition", contrast = c("Condition", "CTd180", "CTd0"))
  histone <- dba.analyze(histone)
  histone
  
  #set working directory to store DiffBind output files
  setwd("output/CUT&Tag/DiffBind")
  
  #save output
  dba.report(histone, contrast=1, method=DBA_DESEQ2, bUsePval=FALSE, bNormalized=TRUE, file=paste(mark, "CTd120vsCTd0_default_all", sep = "_"))
  dba.report(histone, contrast=2, method=DBA_DESEQ2, bUsePval=FALSE, bNormalized=TRUE, file=paste(mark, "CTd180vsCTd0_default_all", sep = "_"))
  
  #PCA plot
  print(mark)
  mypalette=c("green4", "blue", "darkblue")
  dba.plotPCA(histone, DBA_CONDITION, label = DBA_TISSUE, vColors = mypalette, labelSize = 0.5) 
  
  #set working directory back to the original
  setwd("../../../")
  
}

```

### Format DiffBind output csv files to BED files
Output BED files with columns of Chr(V1), Start(V2), End(V3), DiffBindpeak_#(V4), Fold(V5), p.value(V6), FDR(V7)
```{bash}

workingdir="output/CUT&Tag/DiffBind/"

#delete the col names, sort by absolute value in the 7th column (log2 fold change), change scientific notation of chromasome coordinate to normal form, if any
for ifile in $workingdir/*_CTd1[28]0vsCTd0_default_all.csv; do
  sampleName=$(echo "$ifile" | awk -F'/' '{print $NF}' | sed -r 's/.csv//g' | sed -r 's/DBA_//g')
  echo "working on $sampleName..."
  awk 'NR==1{next} NR>1 {print $0}' $ifile | sed -r 's/-([0-9.]+)/\1-/g;' | sed 's/"//g' | sort -t , -k7,7nr | sed -r 's/([0-9.]+)-/-\1/g' | awk -F',' '{OFS="\t"; $2=sprintf("%d", $2); $3=sprintf("%d", $3); print "chr"$1, $2, $3, "DiffBindpeak_"NR, $7, $8, $9}' | sed 's/chrMT/chrM/g' | grep -v "chrKI" | grep -v "chrGL" > $workingdir/${sampleName}.bed
done

```


### peak annotation
#### #K4me3 / K27me3: annotate to promoters using ChIPseeker 
Output txt files with columns of "seqnames", "start", "end", "width", "strand", "V4", "V5", "V6", "V7", "annotation", "geneChr", "geneStart", "geneEnd", "geneLength", "geneStrand", "geneId", "transcriptId", "distanceToTSS", "ENSEMBL", "SYMBOL", "GENENAME" (V4, V5, V6, V7 as from above BED file).
```{r, warning=FALSE, message=FALSE}

workingdir <- "output/CUT&Tag/DiffBind/"

txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

samplefiles <- list.files(path = workingdir, pattern = "K*me3_CTd1[28]0vsCTd0_default_all.bed", full.names=T)
samplefiles <- as.list(samplefiles)
names(samplefiles) <- c("K27me3_CTd120", "K27me3_CTd180", "K4me3_CTd120", "K4me3_CTd180")

#peak annotation with ChIPseeker
peakAnnoList <- lapply(samplefiles, annotatePeak, TxDb=txdb, tssRegion=c(-2000, 2000), verbose=FALSE, annoDb="org.Hs.eg.db")

#plot
plotAnnoBar(peakAnnoList)

#write output 
for (name in names(samplefiles)){
  annot <- data.frame(peakAnnoList[[name]]@anno)
  write.table(annot, file = paste0(workingdir, name, "_diff_peak_annotation_promoter.txt"), sep = "\t", row.names = F, quote = T)
}

```

#### K27ac: annotate to enhancers using bedtools intersect with Genehancer database
Output txt files with columns of chr(V1), element_start(V2), element_end(V3), GHid--GeneHancer element identifier(V4), is_elite(V5), regulatory_element_type(V6), Chr(V7), Start(V8), End(V9), DiffBindpeak_#(V10), Fold (V11), p-value (V12), FDR (V13).
```{bash}

workingdir="output/CUT&Tag/DiffBind/"

#Genehancer database annotation file
Genehancer="data_for_import/CUT&Tag/GeneHancer_AnnotSV_elements_v5.18_elite_enhancer.bed"

inputFiles=("K27ac_CTd120vsCTd0_default_all.bed" "K27ac_CTd180vsCTd0_default_all.bed")

for ifile in "${inputFiles[@]}"
do
  sampleName=$(echo "$ifile" | awk -F'/' '{print $NF}' | sed -r 's/vsCTd0_default_all.bed//g')
  echo $sampleName
  bedtools intersect -wa -wb -a $Genehancer -b $workingdir/$ifile > $workingdir/${sampleName}_diff_peak_annotation_enhancer.txt
done

```

```{r, message = FALSE}
sessionInfo()
```

