---
title: "Initial Analysis of Human Cells out of Shiverer Chimeras"
author: "John Mariani"
date: "12/6/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


```{r, echo = TRUE, message=FALSE, warning=FALSE}
library(Seurat)
library(scPlottingTools)
library(ggplot2)
library(tidyr)
library(dplyr)
library(MAST)
library(plyr)
library(xlsx)
library(patchwork)
library(ggplot2)
library(scales)
library(ggVennDiagram)
library(data.table)
library(slingshot)
library(tradeSeq)
library(tidyr)
library(dplyr)
library(plyr)
library(magrittr)
library(viridis)
library(tidyr)
library(EnhancedVolcano)

`%not in%` <- function(x, table) is.na(match(x, table, nomatch = NA_integer_))

axisTitleSize <- 24
axisTextSize <- 18
labelFont = 18
titleFont = 26
tagSize = 26
fig.pt.size = 2

theme_manuscript <-  theme(axis.text = element_text(size = axisTextSize), 
        axis.title = element_text(size = axisTitleSize), 
        title = element_text(size = titleFont), 
        legend.title = element_text(size = titleFont),
        legend.text = element_text(size = axisTitleSize),
        plot.tag = element_text(size = tagSize))

manuscriptPalette <- c("In Vivo" = "red2", 
                       "In Vitro - GPC Stage" = "#2E30FF",
                       "NPC" = "magenta",
                       "GPC1" = "forestgreen",
                       "GPC2" = "darkorange",
                       "GPC3" = "firebrick2",
                       "GPC4" = "turquoise",
                       "Astrocyte" = "dodgerblue2",
                       "imOL" = "gold",
                       "maOL" = "darkorchid4")

source("Scripts/HelperFunctions.R")




```



## Make invitro only subset
```{r}

invitroInvivo <- readRDS("output/RDS/invitroInvivo.rds")

invitro <- subset(invitroInvivo, subset = stage == "In Vitro - GPC Stage")

DimPlot(invitro, split.by = "cellType")

invitroDE <- subset(invitro, subset = cellType %in% c("NPC", "GPC1", "GPC2", "GPC3", "GPC4"))

DimPlot(invitroDE, split.by = "cellType")

invitroDE <- NormalizeData(invitroDE)


```

## Determine Gene Expresion Fractions and keep those that are present in 10% of any of the in vitro clusters
```{r}

expressionFractions <- DotPlot(invitroDE, assay = "RNA", features = row.names(invitro))$data
names(expressionFractions)

expressionFractionsFilt <- expressionFractions[expressionFractions$pct.exp > 10,]
highFraction <- unique(expressionFractionsFilt$features.plot)

expressionFractionsDF <- pivot_wider(data = expressionFractions, values_from = pct.exp, names_from = id, id_cols = "features.plot")

```

```{r}

# invitro.sca <- makeSCA(invitroDE, highFraction) 
# dim(invitro.sca)

```

## Fit mixed model with fixed cell line, 10x chemistry, and feature complexity effects and a capture random effect
```{r}

# options(mc.cores=8)
# getOption("mc.cores")
# 
# modelMAST <- as.formula(object = "~cellType+line+chemistry+ngeneson+(1|orig.ident)")
# 
# ZLM.invitro <-MAST::zlm(formula = modelMAST, sca = invitro.sca, method='glmer',ebayes = F,
#                                             strictConvergence = FALSE, parallel = T)
# 
# #saveRDS(ZLM.invitro, "output/DE/ZLM.invitro.rds")
# ZLM.invitro <- readRDS("output/DE/ZLM.invitro.rds")
# colnames(ZLM.invitro@coefC)

```

## DE for enrichment in each cell state
```{r}

# options(mc.cores=8)
# getOption("mc.cores")
# 
# #NPC vs rest 
# temp <- colnames(ZLM.invitro@coefC)
# names(temp) <- temp
# temp[] <- c(0,-1/4,-1/4,-1/4,-1/4,0,0,0)
# temp
# 
# runLR(ZLM.invitro, lrContrast = c(0,-1/4,-1/4,-1/4,-1/4,0,0,0),
#       contrast1 = c(1,0,0,0,0,1/2,1/2,0), 
#       contrast0 = c(1,1/4,1/4,1/4,1/4,1/2,1/2,0), FDR = 0.01, logFC = .25, fileName = "NPC.vs.Rest")
# 
# # GPC4 vs rest
# temp[] <- c(0, -1/4, -1/4, -1/4, 1,0,0,0)
# temp
# 
# runLR(ZLM.invitro, lrContrast = c(0, -1/4, -1/4, -1/4, 1,0,0,0),
#       contrast1 = c(1,0,0,0,1,1/2,1/2,0), 
#       contrast0 = c(1,1/4,1/4,1/4,0,1/2,1/2,0), FDR = 0.01, logFC = .25, fileName = "GPC4.vs.Rest")
# 
# 
# # GPC1 vs rest
# temp[] <- c(0, 1, -1/4, -1/4, -1/4,0,0,0)
# temp
# 
# runLR(ZLM.invitro, lrContrast = c(0, 1, -1/4, -1/4, -1/4,0,0,0),
#       contrast1 = c(1,1,0,0,0,1/2,1/2,0), 
#       contrast0 = c(1, 0, 1/4, 1/4, 1/4,1/2,1/2,0), FDR = 0.01, logFC = .25, fileName = "GPC1.vs.Rest")
#  
# # GPC2 vs rest
# temp[] <- c(0, -1/4, 1, -1/4, -1/4,0,0,0)
# temp
# 
# runLR(ZLM.invitro, lrContrast = c(0, -1/4, 1, -1/4, -1/4,0,0,0),
#       contrast1 = c(1,0,1,0,0,1/2,1/2,0), 
#       contrast0 = c(1, 1/4, 0, 1/4, 1/4,1/2,1/2,0), FDR = 0.01, logFC = .25, fileName = "GPC2.vs.Rest")
#  
# #GPC3 vs rest
# temp[] <- c(0, -1/4, -1/4, 1, -1/4,0,0,0)
# temp
# 
# runLR(ZLM.invitro, lrContrast = c(0, -1/4, -1/4, 1, -1/4,0,0,0),
#       contrast1 = c(1,0,0,1,0,1/2,1/2,0), 
#       contrast0 = c(1, 1/4, 1/4, 0, 1/4,1/2,1/2,0), FDR = 0.01, logFC = .25, fileName = "GPC3.vs.Rest")


```



## Reload DE to save time

```{r}

invitroComparisons <- c("GPC4.vs.Rest", "NPC.vs.Rest", "GPC1.vs.Rest", "GPC2.vs.Rest", "GPC3.vs.Rest")

for(i in invitroComparisons){
  temp <- assign(i, read.delim(paste0("output/DE/",i,".txt")))
  print(dim(temp))
  temp$comparison <- i
  temp <- temp[order(temp$logFC, decreasing = T),]
  assign(paste0(i,".sig"), temp[temp$FDR < 0.01 & abs(temp$logFC) > .25,])
}


```



## Expression HM
```{r}

# How many top genes to show in heatmap
howMany <- 11

topHowMany <- GPC4.vs.Rest.sig$gene[1:howMany]
topHowMany <- c(topHowMany, GPC3.vs.Rest.sig[GPC3.vs.Rest.sig$gene %not in% topHowMany,]$gene[1:howMany])
topHowMany <- c(topHowMany, GPC2.vs.Rest.sig[GPC2.vs.Rest.sig$gene %not in% topHowMany,]$gene[1:howMany])
topHowMany <- c(topHowMany, GPC1.vs.Rest.sig[GPC1.vs.Rest.sig$gene %not in% topHowMany & !grepl(pattern = "ENSG", GPC1.vs.Rest.sig$gene),]$gene[1:howMany])
topHowMany <- c(topHowMany, NPC.vs.Rest.sig[NPC.vs.Rest.sig$gene %not in% topHowMany & !grepl(pattern = "ENSG", NPC.vs.Rest.sig$gene),]$gene[1:howMany])


allDE <- rbindlist(list(GPC4.vs.Rest, GPC3.vs.Rest, GPC2.vs.Rest, GPC1.vs.Rest, NPC.vs.Rest))
allDE <- allDE[allDE$gene %in% topHowMany,]
allDE$comparison <- gsub(x = allDE$comparison, pattern = ".vs.Rest", replacement = "")
allDE$comparison <- factor(allDE$comparison, levels = rev(c("NPC", "GPC1", "GPC2", "GPC3","GPC4")))
allDE$gene <- factor(allDE$gene, levels = rev(topHowMany))

allDE$sig <- symnum(allDE$FDR, cutpoints = c(0, 0.0001,
    0.001, 0.01, 0.05, 1), symbols = c("****","***", "**", "*"," "))

allDE$sig <- ifelse(abs(allDE$logFC) < .25, " ", allDE$sig)

expressionHMFig <- ggplot(allDE, aes(y = comparison, x = gene, fill = logFC)) + geom_tile(colour = "black") + scale_fill_gradient2(low = "dodgerblue2", mid = "white", high = "red2", midpoint = 0) + scale_y_discrete(expand = c(0,0)) +theme_bw() + theme_manuscript + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(1, "inch")) + geom_text(aes(label = sig), angle = 90, vjust = 1, size = 12) + guides(fill=guide_colorbar(title="Log2FC Enrichment", title.position = "left", title.theme = element_text(angle = 90, hjust = .5)))

expressionHMFig


           

```


## Integrated Dim Plot
```{r}

dim(invitroInvivo)

integratedDimFig <- DimPlotCustom(invitroInvivo, group.by = "cellType", label = T, pt.size = fig.pt.size) + theme_bw() + theme_manuscript + theme(legend.position = "bottom") + labs(title = "Integrated - 43,142 Cells") + scale_fill_manual(values = manuscriptPalette, aes(legend.title = "CellType")) + guides(fill = guide_legend(override.aes = list(size = 5)))

integratedDimFig

```

## Dim Plot Split by stage
```{r}

stageDimFig <- DimPlotCustom(invitroInvivo, split.by = "stage", group.by = "cellType", ncol = 1, pt.size = fig.pt.size) & theme_bw() & theme_manuscript & NoLegend() & theme(axis.title.y = element_blank())  & scale_fill_manual(values = manuscriptPalette)

table(invitroInvivo$stage)

stageDimFig[[1]] <- stageDimFig[[1]] + labs(title = "In Vivo - 5,337 Cells") 
stageDimFig[[2]] <- stageDimFig[[2]] + xlab("UMAP 1") + labs(title = "In Vitro - 37,805 Cells")
stageDimFig


stageDimFig[[2]] | stageDimFig[[1]]


```

## Stacked Plot Cell Type by Stage
```{r}

StackedCellType <- as.data.frame(table(invitroInvivo$stage, invitroInvivo$cellType))

unique(StackedCellType$Var1)

stackedPlotInvitroInvivo <- ggplot(StackedCellType, aes(fill = Var2, y = Freq, x = Var1))+
  geom_bar(position = "fill", stat = "identity")+
  guides(fill = guide_legend(override.aes = list(size = .6))) + theme_bw() + theme_manuscript + scale_y_continuous(expand = c(0,0))  + labs(y = "Fraction Celltype", x = "Stage", fill = "Cell Type") + theme(legend.key.size = unit(.6, "lines")) + xlab(element_blank()) + NoLegend() & scale_fill_manual(values = manuscriptPalette)

stackedPlotInvitroInvivo 

 
```


## Violin Plots
```{r}


gpcViolins <- (VlnPlot(invitroDE, c("PDGFRA"), pt.size = 0) & coord_flip() & scale_fill_manual(values = manuscriptPalette)& scale_y_continuous(expand = c(0,0), limits = c(0,3.75)) & theme_bw() & theme_manuscript & NoLegend() +  theme(axis.title.y = element_blank(), axis.title.x = element_blank())) /
(VlnPlot(invitroDE, c("NEUROD1"), pt.size = .01) & coord_flip() & scale_fill_manual(values = manuscriptPalette) & scale_y_continuous(expand = c(0,0), limits = c(0,3.75)) & theme_bw() & theme_manuscript & NoLegend() + theme(axis.title.y = element_blank()))




```


## Load in Palantir Data and update seurat metadata
```{r}

palantirMetadata <- read.csv("output/Palantir/palantirInvitroInvivoMetadata.csv", row.names = 1)
palantirMetadata <- palantirMetadata[,c(14:21)]

invitroInvivoMeta <- invitroInvivo@meta.data
invitroInvivoMeta <- merge(invitroInvivoMeta, palantirMetadata, by.x = 0, by.y = 0)
row.names(invitroInvivoMeta) <- invitroInvivoMeta$Row.names
invitroInvivoMeta$Row.names <- NULL

identical(row.names(invitroInvivoMeta), Cells(invitroInvivo))

invitroInvivoMeta <- invitroInvivoMeta[match(Cells(invitroInvivo), row.names(invitroInvivoMeta)),]

identical(row.names(invitroInvivoMeta), Cells(invitroInvivo))

cellLevels <-c("Unselected", "NPC", "GPC1", "GPC2", "GPC3", "GPC4", "imOL", "maOL", "Astrocyte")


invitroInvivoMeta$NPC_branch <- factor(ifelse(invitroInvivoMeta$NPC_branch == T, as.character(invitroInvivoMeta$cellType), "Unselected"), levels = cellLevels)
invitroInvivoMeta$Oligo_branch <- factor(ifelse(invitroInvivoMeta$Oligo_branch == T, as.character(invitroInvivoMeta$cellType), "Unselected"), levels = cellLevels)
invitroInvivoMeta$Astro_branch <- factor(ifelse(invitroInvivoMeta$Astro_branch == T, as.character(invitroInvivoMeta$cellType), "Unselected"), levels = cellLevels)


invitroInvivoMeta$NPC_branch <- factor(invitroInvivoMeta$NPC_branch , levels = cellLevels)
invitroInvivoMeta$Oligo_branch <- factor(invitroInvivoMeta$Oligo_branch , levels = cellLevels)
invitroInvivoMeta$Astro_branch <- factor(invitroInvivoMeta$Astro_branch , levels = cellLevels)


levels(invitroInvivoMeta$NPC_branch)


invitroInvivo@meta.data <- invitroInvivoMeta



```


```{r}


xmin <- min(invitroInvivo@reductions$umap@cell.embeddings[,1])
xmax <- max(invitroInvivo@reductions$umap@cell.embeddings[,1])

ymin <- min(invitroInvivo@reductions$umap@cell.embeddings[,2])
ymax <- max(invitroInvivo@reductions$umap@cell.embeddings[,2])


third <- (DimPlotCustom(subset(invitroInvivo, subset = Oligo_branch != "Unselected"), group.by = "Oligo_branch", label = T, pt.size = fig.pt.size) & scale_fill_manual(values = manuscriptPalette) & ggtitle("Oligodendrocyte Branch") & labs(tag = "G") | 
DimPlotCustom(subset(invitroInvivo, subset = Astro_branch != "Unselected"), group.by = "Astro_branch", label = T, pt.size = fig.pt.size) & scale_fill_manual(values = manuscriptPalette) & ggtitle("Astrocyte Branch") & labs(tag = "I") |
DimPlotCustom(subset(invitroInvivo, subset = NPC_branch != "Unselected"), group.by = "NPC_branch", label = T, pt.size = fig.pt.size) & scale_fill_manual(values = manuscriptPalette) & ggtitle("NPC Branch") & labs(tag = "K")) & theme_bw() & theme_manuscript & NoLegend() & xlim(xmin, xmax) & ylim(ymin, ymax)

third[[2]] <- third[[2]] & theme(axis.title.y = element_blank())

third[[3]] <- third[[3]] & theme(axis.title.y = element_blank())

third


```

```{r}

levels(invitroInvivoMeta$Oligo_branch)

invitroInvivoMeta$Oligo_branch <- factor(invitroInvivoMeta$Oligo_branch, levels = cellLevels)

bottom <- (ggplot(invitroInvivoMeta, aes(x = invitroInvivoMeta$palantir_pseudotime, y = invitroInvivoMeta$Oligo_bp, colour = invitroInvivoMeta$Oligo_branch, group = invitroInvivoMeta$Oligo_branch)) + geom_point(size = 3) + scale_color_manual(values = manuscriptPalette) & labs(tag = "H") |
ggplot(invitroInvivoMeta, aes(x = invitroInvivoMeta$palantir_pseudotime, y = invitroInvivoMeta$Astro_bp, colour = invitroInvivoMeta$Astro_branch)) + geom_point(size = 3) + scale_color_manual(values = manuscriptPalette) & labs(tag = "J") | 
ggplot(invitroInvivoMeta, aes(x = invitroInvivoMeta$palantir_pseudotime, y = invitroInvivoMeta$NPC_bp, colour = invitroInvivoMeta$NPC_branch)) + geom_point(size = 3) + scale_color_manual(values = manuscriptPalette) & labs(tag = "L")) & theme_bw() & theme_manuscript & NoLegend() & labs(x = "Pseudotime", y = "Branch Probability")

bottom[[2]] <- bottom[[2]] & theme(axis.title.y = element_blank())

bottom[[3]] <- bottom[[3]] & theme(axis.title.y = element_blank())

bottom



```

## Piece together main figure
```{r}

integratedDimFig <- integratedDimFig + labs(tag = "A")
stageDimFig[[2]] <- stageDimFig[[2]] + labs(tag = "B")
stageDimFig[[1]] <- stageDimFig[[1]] + labs(tag = "C")
stackedPlotInvitroInvivo <- stackedPlotInvitroInvivo + labs(tag = "D")
gpcViolins[[1]] <- gpcViolins[[1]] + labs(tag = "E")


top <- (integratedDimFig | stageDimFig[[2]] | stageDimFig[[1]] | (stackedPlotInvitroInvivo / gpcViolins))  + plot_layout(widths = c(1,1,1,.45))

second <- (expressionHMFig + labs(tag = "F") + plot_spacer()) + plot_layout(widths = c(1,.001))

(top / second / third / bottom) + plot_layout(heights = c(1,.5,1,.5))

ggsave("output/Figures/Invitro/invitroFig.pdf", width = 30, height = 40)


```
# Supplementary Plots

## Leiden Cluster plot
```{r}

dimLeidenFig <- DimPlotCustom(invitroInvivo, group.by = "leiden_clusters", ncol = 1, pt.size = fig.pt.size, label = T) & theme_bw() & theme_manuscript & NoLegend()  & ggtitle("In Vitro and In Vivo Leiden Clusters")

dimLeidenFig


```


## Split by line dim plot

```{r}

dimLineFig <- DimPlotCustom(invitro, split.by = "line", group.by = "cellType", ncol = 1, pt.size = fig.pt.size, label = T) & theme_bw() & theme_manuscript & NoLegend() & theme(axis.title.y = element_blank()) & scale_fill_manual(values = manuscriptPalette)

dimLineFig

table(invitro$line)

dimLineFig[[1]] <- dimLineFig[[1]] + labs(title = "iPSC (C27) - 16,808 In Vitro Cells") + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
dimLineFig[[2]] <- dimLineFig[[2]] + xlab("UMAP 1") + labs(title = "ESC (WA09) - 20,997 In Vitro Cells")
dimLineFig





```


## Sort
```{r}

StackedSort <- as.data.frame(table(invitro$sort, invitro$cellType))
StackedSort$Var1 <- relevel(StackedSort$Var1, "DAPI-")

unique(StackedSort$Var1)

StackedSortPlot <- ggplot(StackedSort, aes(fill = Var2, y = Freq, x = Var1))+
  geom_bar(position = "fill", stat = "identity")+
  guides(fill = guide_legend(override.aes = list(size = .6))) + theme_bw() + theme_manuscript + scale_y_continuous(expand = c(0,0)) + ylab("Percent Identity") + labs(y = "Percent", x = "Stage", fill = "Cell Type", tag = "C") + theme(legend.key.size = unit(.6, "lines")) + xlab(element_blank()) + NoLegend() & scale_fill_manual(values = manuscriptPalette)  & ggtitle("Sort Cell Type Fractions")

StackedSortPlot

```

## Cell Cycle Scoring

```{r}

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

#MLF1IP
s.genes[s.genes %not in% invitro@assays$RNA@counts@Dimnames[[1]]] <- "CENPU"
s.genes[s.genes %not in% invitro@assays$RNA@counts@Dimnames[[1]]]
#FAM64A HN1
g2m.genes[g2m.genes %not in% invitro@assays$RNA@counts@Dimnames[[1]]] <- c("PIMREG", "JPT1")
g2m.genes[g2m.genes %not in% invitro@assays$RNA@counts@Dimnames[[1]]]

invitro <- CellCycleScoring(invitro, s.features = s.genes, g2m.features = g2m.genes, set.ident = F)


DimPlotCustom(invitro, split.by = "Phase", label = T, group.by = "cellType") 



cellCycleVlnGG <- VlnPlot(invitro, c("G2M.Score", "S.Score"), pt.size = 0, idents = c("NPC", "GPC1", "GPC2", "GPC3", "GPC4")) & theme_bw() & theme_manuscript & NoLegend() & theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

cellCycleVlnGG[[1]] <- cellCycleVlnGG[[1]] + ggtitle("G2M Phase Score")
cellCycleVlnGG[[2]] <- cellCycleVlnGG[[2]] + ggtitle("S Phase Score")

cellCycleVlnGG


color0 = "grey60"
colPalette = c("dodgerblue2", 
        "gold", "red2")

cellCycleFeatureGG <- FeaturePlot(invitro, c("G2M.Score", "S.Score"), order = T, pt.size = fig.pt.size) & scale_colour_gradientn(colors = c(color0, colorRampPalette(colPalette)(100))) & theme_bw() & theme_manuscript & theme(legend.position = "bottom", legend.key.width = unit(.5, "inch")) & xlab("UMAP 1") & ylab("UMAP 2")

cellCycleFeatureGG[[1]] <- cellCycleFeatureGG[[1]] + ggtitle ("G2M Phase Score")
cellCycleFeatureGG[[2]] <- cellCycleFeatureGG[[2]] + ggtitle ("S Phase Score")


cellCycleFeatureGG

```

## Mature NPC Feature Plots
```{r}

npcFeaturePlots <- FeaturePlotCustom(invitro, c("SYT1", "MEIS2", "BCL11B"), plotLegend = "shared", pt.size = fig.pt.size)   & theme_bw() & theme_manuscript & theme(legend.position = "bottom", legend.key.width = unit(.5, "inch"))

npcFeaturePlots[[1]] <- npcFeaturePlots[[1]] + ylab("UMAP 2")
npcFeaturePlots <- npcFeaturePlots & xlab("UMAP 1")

npcFeaturePlots

```

## Supplemental Figure
```{r}

dimLeidenFig <- dimLeidenFig + labs(tag = "A")
dimLineFig[[1]] <- dimLineFig[[1]] + labs(tag = "B")
dimLineFig[[2]] <- dimLineFig[[2]] + labs(tag = "C")
StackedSortPlot <- StackedSortPlot + labs(tag = "D")
cellCycleFeatureGG[[1]] <- cellCycleFeatureGG[[1]] + labs(tag = "E")
cellCycleVlnGG[[1]] <- cellCycleVlnGG[[1]] + labs(tag = "F")
npcFeaturePlots[[1]] <- npcFeaturePlots[[1]] + labs(tag = "G")


top <- dimLeidenFig | dimLineFig[[1]] | dimLineFig[[2]] | StackedSortPlot

middle <- cellCycleFeatureGG | cellCycleVlnGG

bottom <- npcFeaturePlots

top / middle / bottom

ggsave("output/Figures/Invitro/invitroSuppFig.pdf", width = 30, height = 40)



```


```{r}

sessionInfo()

```