---
title: "Differential NicheNet Analysis of in vivo and in vitro hGPCs"
author: "John Mariani"
date: "3/6/2023"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

## Load in Libraries

```{r}



library(dplyr)
library(Seurat)
library(devtools)
library(patchwork)
library(dplyr)
library(SeuratDisk)
library(biomaRt)
library(Matrix)
library(nichenetr)
library(RColorBrewer)
library(scPlottingTools)
library(data.table)

options(future.globals.maxSize = 16000 * 1024^2)

```

## Read in Human and Mouse Counts

```{r}

source("Scripts/HelperFunctions.r")

dualSpecies <- readRDS("output/RDS/finalNiche.rds")

```


## Read in NicheNetV2 Databases
```{r}

ligand_target_matrix = readRDS("data_for_import/NicheNet/ligand_target_matrix_nsga2r_final.rds")
lr_network = readRDS("data_for_import/NicheNet/lr_network_human_21122021.rds")
lr_network = lr_network %>% mutate(bonafide = ! database %in% c("ppi_prediction","ppi_prediction_go"))
lr_network = lr_network %>% dplyr::rename(ligand = from, receptor = to) %>% distinct(ligand, receptor, bonafide)

weighted_networks = readRDS("data_for_import/NicheNet/weighted_networks_nsga2r_final.rds")

```

## Make counts to edit
```{r}

dualSpecies$stage2 <- dualSpecies$stage
dualSpecies$stage2 <- gsub(x = dualSpecies$stage2, pattern = "In Vitro - GPC Stage", replacement = "In Vitro")
unique(dualSpecies$stage2)

dualSpecies$cellStage <- paste0(dualSpecies$stage2, " ", dualSpecies$cellType)
dualSpecies$cellStage <- gsub(x = dualSpecies$cellStage, pattern = "Immature Oligodendrocyte", replacement = "imOL")
dualSpecies$cellStage <- gsub(x = dualSpecies$cellStage, pattern = "Mature Oligodendrocyte", replacement = "maOL")

table(dualSpecies$cellStage)


```

## Showing ligand differences between in vitro cell states
```{r}

ligands <- unique(lr_network$ligand)

GPC1 <- read.delim("output/DE/GPC1.vs.Rest.sig.txt")
GPC2 <- read.delim("output/DE/GPC2.vs.Rest.sig.txt")
GPC3 <- read.delim("output/DE/GPC3.vs.Rest.sig.txt")
GPC4 <- read.delim("output/DE/GPC4.vs.Rest.sig.txt")
NPC <- read.delim("output/DE/NPC.vs.Rest.sig.txt")

invitroStateMarkers <- rbind(GPC1, GPC2, GPC3, GPC4, NPC)
invitroStateLigands <- invitroStateMarkers[invitroStateMarkers$gene %in% ligands,]
invitroStateLigands <- invitroStateLigands[!duplicated(invitroStateLigands$gene),]


```



```{r}

niches = list(
    "In Vivo" = list(
      "sender" = c("In Vivo Astrocyte","In Vivo imOL", "In Vivo maOL", "In Vivo Mouse Astrocyte", "In Vivo Mouse Endothelial", "In Vivo Mouse Ependymal", "In Vivo Mouse Macrophage", "In Vivo Mouse GPC", "In Vivo Mouse imOL", "In Vivo Mouse maOL", "In Vivo Mouse Microglia", "In Vivo Mouse NPC", "In Vivo Mouse Pericyte"),
      "receiver" = c("In Vivo GPC4")),
    "In Vitro" = list(
      "sender" = c("In Vitro GPC1","In Vitro GPC2", "In Vitro GPC3", "In Vitro NPC"),
      "receiver" = c("In Vitro GPC4"))
  )



Idents(dualSpecies) <- dualSpecies$cellStage

one2oneGenes <- read.csv("data_for_import/NicheNet/one2oneGenes.csv")

#DE_sender = calculate_niche_de(seurat_obj = dualSpecies %>% subset(features = lr_network$ligand %>% intersect(one2oneGenes$x)), niches = niches, type = "sender", assay_oi = "RNA") # only ligands important for sender cell types


#write.csv(DE_sender, "output/NicheNet/DE_sender.csv", quote = F, row.names = F)

DE_sender <- read.csv("output/NicheNet/DE_sender.csv")
DE_sender$sender <- gsub(x = DE_sender$sender, pattern = "Immature Oligodendrocyte", replacement = "imOL")
DE_sender$sender_other_niche <- gsub(x = DE_sender$sender_other_niche, pattern = "Immature Oligodendrocyte", replacement = "imOL")
DE_sender$sender <- gsub(x = DE_sender$sender, pattern = "Mature Oligodendrocyte", replacement = "maOL")
DE_sender$sender_other_niche <- gsub(x = DE_sender$sender_other_niche, pattern = "Mature Oligodendrocyte", replacement = "maOL")
DE_sender$sender <- gsub(x = DE_sender$sender, pattern = "Immature Oligo", replacement = "imOL")
DE_sender$sender_other_niche <- gsub(x = DE_sender$sender_other_niche, pattern = "Immature Oligo", replacement = "imOL")
DE_sender$sender <- gsub(x = DE_sender$sender, pattern = "Mature Oligo", replacement = "maOL")
DE_sender$sender_other_niche <- gsub(x = DE_sender$sender_other_niche, pattern = "Mature Oligo", replacement = "maOL")


deGPC4 <- read.delim("output/DE/Invivo.vs.Invitro.GPC4.txt")
deGPC4.sig <- read.delim("output/DE/Invivo.vs.Invitro.GPC4.sig.txt")
deGPC4 <- deGPC4[complete.cases(deGPC4),]
deGPC4 <- rbind(deGPC4, deGPC4.sig[deGPC4.sig$gene %not in% deGPC4$gene,])

#Format Prior Differential Expression for receiver cells

pct.expr <- DotPlot(dualSpecies, features = deGPC4$gene, idents = c("In Vivo GPC4", "In Vitro GPC4"), group.by = "cellStage")$data
pct.expr.invivo <- pct.expr[pct.expr$id == "In Vivo GPC4",]
pct.expr.invitro <- pct.expr[pct.expr$id == "In Vitro GPC4",]


DE_receiver_invivo <- data.frame(gene = deGPC4$gene, p_val = deGPC4$FDR, avg_log2FC = deGPC4$logFC, pct.1 = pct.expr.invivo$pct.exp, pct.2 = pct.expr.invitro$pct.exp, p_val_adj = deGPC4$FDR, receiver = "In Vivo GPC4", receiver_other_niche = "In Vitro GPC4")

DE_receiver_invitro <- data.frame(gene = deGPC4$gene, p_val = deGPC4$FDR, avg_log2FC = (deGPC4$logFC*-1), pct.1 = pct.expr.invitro$pct.exp, pct.2 = pct.expr.invivo$pct.exp, p_val_adj = deGPC4$FDR, receiver = "In Vitro GPC4", receiver_other_niche = "In Vivo GPC4")


DE_receiver <- rbind(DE_receiver_invivo, DE_receiver_invitro)


DE_receiver_receptors <- DE_receiver[DE_receiver$gene %in% lr_network$receptor,]

#DE_receiver = calculate_niche_de(seurat_obj = dualSpecies %>% subset(features = lr_network$receptor %>% unique()), niches = niches, type = "receiver", assay_oi = "RNA") # only receptors now, later on: DE analysis to find targets

DE_sender = DE_sender %>% mutate(avg_log2FC = ifelse(avg_log2FC == Inf, max(avg_log2FC[is.finite(avg_log2FC)]), ifelse(avg_log2FC == -Inf, min(avg_log2FC[is.finite(avg_log2FC)]), avg_log2FC)))

DE_receiver_receptors = DE_receiver_receptors %>% mutate(avg_log2FC = ifelse(avg_log2FC == Inf, max(avg_log2FC[is.finite(avg_log2FC)]), ifelse(avg_log2FC == -Inf, min(avg_log2FC[is.finite(avg_log2FC)]), avg_log2FC)))




```

```{r}

expression_pct = 10
DE_sender_processed = process_niche_de(DE_table = DE_sender, niches = niches, expression_pct = expression_pct, type = "sender")
DE_receiver_processed = process_niche_de(DE_table = DE_receiver_receptors, niches = niches, expression_pct = expression_pct, type = "receiver")

specificity_score_LR_pairs = "min_lfc"
DE_sender_receiver = combine_sender_receiver_de(DE_sender_processed, DE_receiver_processed, lr_network, specificity_score = specificity_score_LR_pairs)



```


```{r}

lfc_cutoff = 0.25 # recommended for 10x as min_lfc cutoff. 
specificity_score_targets = "min_lfc"

#DE_receiver_targets = calculate_niche_de_targets(seurat_obj = dualSpecies, niches = niches, lfc_cutoff = lfc_cutoff, expression_pct = expression_pct, assay_oi = "RNA") 

DE_receiver_processed_targets = process_receiver_target_de(DE_receiver = DE_receiver, niches = niches, expression_pct = expression_pct, specificity_score = specificity_score_targets)
  
background = DE_receiver_processed_targets  %>% pull(target) %>% unique()

geneset_InVivo = DE_receiver_processed_targets %>% filter(receiver == niches$`In Vivo`$receiver & target_score >= lfc_cutoff & target_significant == 1 & target_present == 1) %>% pull(target) %>% unique()


geneset_Invitro = DE_receiver_processed_targets %>% filter(receiver == niches$`In Vitro`$receiver & target_score >= lfc_cutoff & target_significant == 1 & target_present == 1) %>% pull(target) %>% unique()

geneset_InVivo %>% setdiff(rownames(ligand_target_matrix))
geneset_Invitro %>% setdiff(rownames(ligand_target_matrix))



```

```{r}

top_n_target = 1000

niche_geneset_list = list(
    "In_Vivo_niche" = list(
      "receiver" = "In Vivo GPC4",
      "geneset" = geneset_InVivo,
      "background" = background),
    "In_Vitro_niche" = list(
      "receiver" = "In Vitro GPC4",
      "geneset" = geneset_Invitro ,
      "background" = background))

ligand_activities_targets = get_ligand_activities_targets(niche_geneset_list = niche_geneset_list, ligand_target_matrix = ligand_target_matrix, top_n_target = top_n_target)

write.table(ligand_activities_targets, "output/NicheNet/ligand_activities_targets.txt", sep = "\t", row.names = F, quote = F)


```

```{r}

features_oi = union(lr_network$ligand, lr_network$receptor) %>% union(ligand_activities_targets$target) %>% setdiff(NA)
  
dotplot = suppressWarnings(Seurat::DotPlot(dualSpecies %>% subset(idents = niches %>% unlist() %>% unique()), features = features_oi, assay = "RNA"))

exprs_tbl = dotplot$data %>% as_tibble()
exprs_tbl = exprs_tbl %>% dplyr::rename(celltype = id, gene = features.plot, expression = avg.exp, expression_scaled = avg.exp.scaled, fraction = pct.exp) %>%
    mutate(fraction = fraction/100) %>% as_tibble() %>% dplyr::select(celltype, gene, expression, expression_scaled, fraction) %>% distinct() %>% arrange(gene) %>% mutate(gene = as.character(gene))
  
exprs_tbl_ligand = exprs_tbl %>% filter(gene %in% lr_network$ligand) %>% dplyr::rename(sender = celltype, ligand = gene, ligand_expression = expression, ligand_expression_scaled = expression_scaled, ligand_fraction = fraction) 

exprs_tbl_receptor = exprs_tbl %>% filter(gene %in% lr_network$receptor) %>% dplyr::rename(receiver = celltype, receptor = gene, receptor_expression = expression, receptor_expression_scaled = expression_scaled, receptor_fraction = fraction)

exprs_tbl_target = exprs_tbl %>% filter(gene %in% ligand_activities_targets$target) %>% dplyr::rename(receiver = celltype, target = gene, target_expression = expression, target_expression_scaled = expression_scaled, target_fraction = fraction)

exprs_tbl_ligand = exprs_tbl_ligand %>%  mutate(scaled_ligand_expression_scaled = scale_quantile_adapted(ligand_expression_scaled)) %>% mutate(ligand_fraction_adapted = ligand_fraction) %>% mutate_cond(ligand_fraction >= expression_pct, ligand_fraction_adapted = expression_pct)  %>% mutate(scaled_ligand_fraction_adapted = scale_quantile_adapted(ligand_fraction_adapted))

exprs_tbl_receptor = exprs_tbl_receptor %>% mutate(scaled_receptor_expression_scaled = scale_quantile_adapted(receptor_expression_scaled))  %>% mutate(receptor_fraction_adapted = receptor_fraction) %>% mutate_cond(receptor_fraction >= expression_pct, receptor_fraction_adapted = expression_pct)  %>% mutate(scaled_receptor_fraction_adapted = scale_quantile_adapted(receptor_fraction_adapted))

```


```{r}

exprs_sender_receiver = lr_network %>% 
  inner_join(exprs_tbl_ligand, by = c("ligand")) %>% 
  inner_join(exprs_tbl_receptor, by = c("receptor")) %>% inner_join(DE_sender_receiver %>% distinct(niche, sender, receiver))
  
ligand_scaled_receptor_expression_fraction_df = exprs_sender_receiver %>% group_by(ligand, receiver) %>% mutate(rank_receptor_expression = dense_rank(receptor_expression), rank_receptor_fraction  = dense_rank(receptor_fraction)) %>% mutate(ligand_scaled_receptor_expression_fraction = 0.5*( (rank_receptor_fraction / max(rank_receptor_fraction)) + ((rank_receptor_expression / max(rank_receptor_expression))) ) )  %>% distinct(ligand, receptor, receiver, ligand_scaled_receptor_expression_fraction, bonafide) %>% distinct() %>% ungroup() 

# No Spatial info
spatial_info = tibble(celltype_region_oi = NA, celltype_other_region = NA) %>% mutate(niche =  niches %>% names() %>% head(1), celltype_type = "sender")


```

## No Spaital info
```{r}


include_spatial_info_sender = F # if not spatial info to include: put this to false 
include_spatial_info_receiver = FALSE # if spatial info to include: put this to true 


if(include_spatial_info_sender == FALSE & include_spatial_info_receiver == FALSE){
    spatial_info = tibble(celltype_region_oi = NA, celltype_other_region = NA) %>% mutate(niche =  niches %>% names() %>% head(1), celltype_type = "sender")
} 


if(include_spatial_info_sender == TRUE){
  sender_spatial_DE = calculate_spatial_DE(seurat_obj = seurat_obj %>% subset(features = lr_network$ligand %>% unique()), spatial_info = spatial_info %>% filter(celltype_type == "sender"), assay_oi = assay_oi)
  sender_spatial_DE_processed = process_spatial_de(DE_table = sender_spatial_DE, type = "sender", lr_network = lr_network, expression_pct = expression_pct, specificity_score = specificity_score_spatial)

  # add a neutral spatial score for sender celltypes in which the spatial is not known / not of importance
  sender_spatial_DE_others = get_non_spatial_de(niches = niches, spatial_info = spatial_info, type = "sender", lr_network = lr_network)
  sender_spatial_DE_processed = sender_spatial_DE_processed %>% bind_rows(sender_spatial_DE_others)

  sender_spatial_DE_processed = sender_spatial_DE_processed %>% mutate(scaled_ligand_score_spatial = scale_quantile_adapted(ligand_score_spatial))

} else {
  # # add a neutral spatial score for all sender celltypes (for none of them, spatial is relevant in this case)
  sender_spatial_DE_processed = get_non_spatial_de(niches = niches, spatial_info = spatial_info, type = "sender", lr_network = lr_network)
  sender_spatial_DE_processed = sender_spatial_DE_processed %>% mutate(scaled_ligand_score_spatial = scale_quantile_adapted(ligand_score_spatial))  

}

if(include_spatial_info_receiver == TRUE){
  receiver_spatial_DE = calculate_spatial_DE(seurat_obj = seurat_obj %>% subset(features = lr_network$receptor %>% unique()), spatial_info = spatial_info %>% filter(celltype_type == "receiver"), assay_oi = assay_oi)
  receiver_spatial_DE_processed = process_spatial_de(DE_table = receiver_spatial_DE, type = "receiver", lr_network = lr_network, expression_pct = expression_pct, specificity_score = specificity_score_spatial)

  # add a neutral spatial score for receiver celltypes in which the spatial is not known / not of importance
  receiver_spatial_DE_others = get_non_spatial_de(niches = niches, spatial_info = spatial_info, type = "receiver", lr_network = lr_network)
  receiver_spatial_DE_processed = receiver_spatial_DE_processed %>% bind_rows(receiver_spatial_DE_others)

  receiver_spatial_DE_processed = receiver_spatial_DE_processed %>% mutate(scaled_receptor_score_spatial = scale_quantile_adapted(receptor_score_spatial))

} else {
    # # add a neutral spatial score for all receiver celltypes (for none of them, spatial is relevant in this case)
  receiver_spatial_DE_processed = get_non_spatial_de(niches = niches, spatial_info = spatial_info, type = "receiver", lr_network = lr_network)
  receiver_spatial_DE_processed = receiver_spatial_DE_processed %>% mutate(scaled_receptor_score_spatial = scale_quantile_adapted(receptor_score_spatial))
}



```


```{r}

prioritizing_weights = c("scaled_ligand_score" = 5,
                         "scaled_ligand_expression_scaled" = 1,
                         "ligand_fraction" = 1,
                         "scaled_ligand_score_spatial" = 0, 
                         "scaled_receptor_score" = 0.5,
                         "scaled_receptor_expression_scaled" = 0.5,
                          "receptor_fraction" = 1, 
                         "ligand_scaled_receptor_expression_fraction" = 1,
                         "scaled_receptor_score_spatial" = 0,
                         "scaled_activity" = 0,
                         "scaled_activity_normalized" = 1,
                         "bona_fide" = 1)


output = list(DE_sender_receiver = DE_sender_receiver, ligand_scaled_receptor_expression_fraction_df = ligand_scaled_receptor_expression_fraction_df, sender_spatial_DE_processed = sender_spatial_DE_processed, receiver_spatial_DE_processed = receiver_spatial_DE_processed,
         ligand_activities_targets = ligand_activities_targets, DE_receiver_processed_targets = DE_receiver_processed_targets, exprs_tbl_ligand = exprs_tbl_ligand,  exprs_tbl_receptor = exprs_tbl_receptor, exprs_tbl_target = exprs_tbl_target)

prioritization_tables = get_prioritization_tables(output, prioritizing_weights)

prioritization_tables$prioritization_tbl_ligand_receptor %>% filter(receiver == niches[[1]]$receiver) %>% head(10)
prioritization_tables$prioritization_tbl_ligand_receptor %>% filter(receiver == niches[[2]]$receiver) %>% head(10)




prioritization_tables$prioritization_tbl_ligand_receptor = prioritization_tables$prioritization_tbl_ligand_receptor %>% mutate(receiver = factor(receiver, levels = c("In Vivo GPC4", "In Vitro GPC4")), niche = factor(niche, levels = c("In Vivo","In Vitro"))) 

prioritization_tables$prioritization_tbl_ligand_target = prioritization_tables$prioritization_tbl_ligand_target %>% mutate(receiver = factor(receiver, levels = c("In Vivo GPC4", "In Vitro GPC4")), niche = factor(niche, levels = c("In Vivo", "In Vitro"))) 



```

## Ligand Activities for picking targets

```{r}


ligandActivities <- ligand_activities_targets[!duplicated(ligand_activities_targets[c("ligand", "receiver")]),]
ligandActivities <- ligandActivities[,-c(4:5)]

ligandActivitiesInVivo <- ligandActivities[ligandActivities$receiver == "In Vivo GPC4",]
ligandActivitiesInVitro <- ligandActivities[ligandActivities$receiver == "In Vitro GPC4",]

ligandActivitiesInVivo <- ligandActivitiesInVivo[,-4]
ligandActivitiesInVitro <- ligandActivitiesInVitro[,-4]


names(ligandActivitiesInVivo)[2:6] <- paste(names(ligandActivitiesInVivo)[2:6], "InVivo", sep = "_")
names(ligandActivitiesInVitro)[2:6] <- paste(names(ligandActivitiesInVitro)[2:6], "InVitro", sep = "_")

ligandActivities <- merge(ligandActivitiesInVivo, ligandActivitiesInVitro, by.x = 1, by.y = 1)
ligandActivities <- ligandActivities[,c(1,2,7,3,8,4,9,5,10,6,11)]

ligandActivities$activity_normalized_diff <- ligandActivities$activity_normalized_InVivo - ligandActivities$activity_normalized_InVitro
ligandActivities$activity_normalized_div <- ligandActivities$activity_normalized_InVivo / ligandActivities$activity_normalized_InVitro


# ligandActivities$aupr_corrected_diff <- ligandActivities$aupr_corrected_InVivo - ligandActivities$aupr_corrected_InVitro
# ligandActivities$aupr_corrected_div <- ligandActivities$aupr_corrected_InVivo / ligandActivities$aupr_corrected_InVitro
# 
# ligandActivities$activity_diff <- ligandActivities$activity_InVivo - ligandActivities$activity_InVitro
# ligandActivities$activity_div <- ligandActivities$activity_InVivo / ligandActivities$activity_InVitro
# 
# 
# ligandActivities$scaled_activity_normalized_diff <- ligandActivities$scaled_activity_normalized_InVivo - ligandActivities$scaled_activity_normalized_InVitro
# ligandActivities$scaled_activity_normalized_div <- ligandActivities$scaled_activity_normalized_InVivo / ligandActivities$scaled_activity_normalized_InVitro

ligandActivities <- ligandActivities[order(ligandActivities$activity_normalized_diff, decreasing = T),]

pri <- prioritization_tables$prioritization_tbl_ligand_receptor

priInvivo <- pri[pri$niche == "In Vivo",]
priInvivo <- priInvivo[priInvivo$receptor_fraction > .1,]
priInvivo <- priInvivo[priInvivo$ligand %in% ligandActivities[ligandActivities$activity_normalized_diff > 0.6,]$ligand,]
priInvivo <- priInvivo[order(priInvivo$ligand_score, decreasing = T),]
#priInvivo <- priInvivo[priInvivo$receptor %in% deGPC4[deGPC4$logFC > 0,]$gene,] 

ligandActivitiesInvivoReceptor <- ligandActivities[ligandActivities$ligand %in% priInvivo$ligand,]
ligandActivitiesInvivoReceptor <- ligandActivitiesInvivoReceptor[ligandActivitiesInvivoReceptor$activity_normalized_diff > 0.6,]

# Differentially regulated in vitro ligands
ligandActivitiesInvivoReceptor[ligandActivitiesInvivoReceptor$ligand %in% invitroStateLigands$gene,]


```


## Ligand Activity and Expression for plots
```{r}

plotting_tbl <- exprs_tbl

plotting_tbl$stage <- ifelse(grepl("Vivo",plotting_tbl$celltype),"In Vivo","In Vitro")
plotting_tbl$species <- ifelse(grepl("Mouse",plotting_tbl$celltype),"Mouse","Human")
plotting_tbl$ct <- gsub(pattern = "In Vitro ", replacement = "", plotting_tbl$celltype)
plotting_tbl$ct <- gsub(pattern = "In Vivo ", replacement = "", plotting_tbl$ct)
plotting_tbl$ct <- gsub(pattern = "Mouse ", replacement = "", plotting_tbl$ct)
unique(plotting_tbl$ct)

temp <- unique(plotting_tbl$ct)
temp
names(temp) <- c("imOL", "maOL", "Astrocyte", "GPC1", "GPC2", "NPC", "GPC3", "GPC", "Microglia", "Ependymal", "Endothelial", "Pericyte", "Macrophage", "GPC4")

temp
plotting_tbl$ctRenamed <- plyr::mapvalues(plotting_tbl$ct, from = temp, to = names(temp))

unique(plotting_tbl$ctRenamed)

plotting_tbl$ctRenamed <- factor(plotting_tbl$ctRenamed, levels = c("GPC1", "GPC2", "GPC3", "GPC4", "GPC", "imOL", "maOL", "Astrocyte", "Microglia", "Macrophage", "NPC", "Ependymal", "Endothelial", "Pericyte"))

write.table(plotting_tbl, "output/NicheNet/plotting_tbl.txt", row.names = F, quote = F, sep = "\t")

ligandsCurated <- c("PDGFA", "OSM", "IL1A", "CSF1", "EGF", "TNF", "EBI3", "IL1B", "NOG", "SLITRK1", "TNFSF10", "GHRH", "TGFB1", "GAS6", "ICAM1", "EDN1", "CCN2", "DSCAM", "MMP13", "ADM", "APOE", "OCLN", "HBEGF", "FGF1")

#ligandsCurated <- ligandActivitiesInvivoReceptor$ligand[81:95]





# ligandsActivity <- c("SLITRK1", "PDGFA", "OSM", "LRRC4", "IL1A", "IFNK", "EGF", "EBI3", "CSF1", "CMTM8", "CLCF1",
#                      "TNFSF10", "TNF", "NOG", "GAS6", "DSCAM", "DKK3", "BDNF")
# 
# ligandsScore <- c("APOE", "CCN2", "DKK3", "EBI3", "GAS6", "ICAM1", "IL1A", "OCLN", "OSM", "PDGFA", "PGF", "TNFSF10", "FGF1",)
# 
# ligandsCurated <- unique(c(ligandsActivity, ligandsScore))




```

```{r}

ligandOrder <- rev(ligandActivities[ligandActivities$ligand %in% ligandsCurated,]$ligand)
ligandPlot <- plotting_tbl[plotting_tbl$gene %in% ligandsCurated,]
ligandPlot$gene <- factor(ligandPlot$gene, levels = ligandOrder)

activityPlot <- ligandActivities[ligandActivities$ligand %in% ligandsCurated,]
activityPlot$ligand <- factor(activityPlot$ligand, levels = ligandOrder)



```




## Ligand Activity and Expression plot
```{r}

ligandActivityGG <- ggplot(activityPlot, aes(x = "In Vivo", y = ligand, fill = scale(activity_normalized_diff))) + 
  geom_tile(colour = "black") + 
  viridis::scale_fill_viridis(option = "D", expand = c(0,0)) + 
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  theme_bw() +
  theme_manuscript + 
  theme(axis.title = element_blank(), legend.title = element_blank(), legend.position = "left", axis.text.y = element_blank()) +
  labs(title = "Ligand Activity", tag = "C")
  

ligandExpressionGG <- ggplot(ligandPlot, aes(x = ctRenamed, y = gene, colour = expression_scaled, size = ifelse(fraction==0, NA, fraction))) + 
  geom_point() + 
  viridis::scale_color_viridis(option = "D") + 
  facet_grid(rows = ~species + stage, scales = "free", space = "free") + 
  theme_bw() +
  theme_manuscript + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), legend.position = "bottom", axis.title = element_blank(), legend.title = element_blank()) +
  scale_size(range = c(0,4)) +
  labs(title = "Ligand Expression", tag = "D")


(ligandActivityGG | ligandExpressionGG) + plot_layout(widths = c(.5,10))

```


## Circos
```{r}

curatedCircos <- c("PDGFA", "IL1A", "OSM", "EBI3", "TGFB1", "TNF", "MMP13", "GAS6", "HBEGF", "ICAM1", "IL1B", "APOE", "EGF", "CSF1", "NOG", "ADM", "TNFSF10", "OCLN", "EDN1", "ADM")

curatedCircos <- c("PDGFA", "OSM", "IL1A", "EBI3", "EGF", "CSF1", "TGFB1", "TNF", "ICAM1", "OCLN")

tempPri <- prioritization_tables$prioritization_tbl_ligand_receptor
tempPri <- tempPri[tempPri$sender %not in% c("In Vivo Mouse NPC", "In Vivo imOL", "In Vivo maOL", "In Vivo Mouse GPC", "In Vivo Mouse Ependymal", "In Vivo Astrocyte"),]

tempPri <- tempPri[tempPri$niche == "In Vivo",]
tempPri <- tempPri[tempPri$receptor_fraction > .1,]


top_ligand_receptor_niche_df = tempPri %>% dplyr::select(niche, sender, receiver, ligand, receptor, prioritization_score) %>% group_by(ligand, receptor) %>% top_n(1, prioritization_score) %>% ungroup() %>% dplyr::select(ligand, receptor, niche) 


prioritized_tbl_oi = tempPri %>% dplyr::filter(ligand %in% curatedCircos) %>% dplyr::select(niche, sender, receiver, ligand,  receptor, ligand_receptor, prioritization_score) %>% distinct() %>% inner_join(top_ligand_receptor_niche_df) %>% group_by(ligand) %>% dplyr::filter(receiver == "In Vivo GPC4") %>% top_n(2, prioritization_score) %>% ungroup() 

unique(prioritized_tbl_oi$sender)

colors_sender <- c("#00BF7D","#D89000", "#E76BF3", "#00B0F6", "#00BFC4", "#39B600", "#A3A500")
names(colors_sender) <- unique(prioritized_tbl_oi$sender)

colors_receiver <- "turquoise"
names(colors_receiver) <- unique(prioritized_tbl_oi$receiver)


circos_output = make_circos_lr(prioritized_tbl_oi, colors_sender, colors_receiver)

unique(prioritized_tbl_oi$sender)




circos_output = make_circos_lr(prioritized_tbl_oi, colors_sender, colors_receiver)


```
## Receptors for Supp Table
```{r}

receptorSup <- pri
receptorSup <- receptorSup[,c(2,5,6,18,20)]
receptorSup <- receptorSup[!duplicated(receptorSup),]
receptorSup <- receptorSup[order(receptorSup$receptor, receptorSup$ligand),]

write.table(receptorSup, "output/NicheNet/receptorSuppTable.txt", quote = F, row.names = F, sep = "\t")

```


## Receptor Expression for plotting

```{r}

##
tempDF <- pri[pri$ligand %in% ligandsCurated,]
length(unique(tempDF$receptor))
tempDF <- tempDF[,c(1, 4:6, 18:20)]
tempDF <- tempDF[!duplicated(tempDF),]
unique(tempDF$receptor)

tempDF <- tempDF[tempDF$receptor %in% priInvivo$receptor,]

tempDF$niche <- factor(tempDF$niche, levels = c("In Vivo", "In Vitro"))



#activityPlot <- activityPlot[activityPlot$ligand %not in% ligandRemove,]
#ligandPlot <- ligandPlot[ligandPlot$gene %not in% ligandRemove,]

tempDF$ligand_receptor <- factor(tempDF$ligand_receptor, levels = unique(tempDF[order(tempDF$ligand, tempDF$receptor),]$ligand_receptor))

tempDF$receptor_fraction <- tempDF$receptor_fraction * 100


```

## Receptor Plot
```{r}

receptorGG <- ggplot(tempDF, aes(y = niche, x = ligand_receptor, fill = receptor_expression_scaled, size = ifelse(receptor_fraction==0, NA, receptor_fraction))) + 
  geom_point(color = "black", pch = 21) + 
  theme_bw() +
  theme_manuscript + 
  scale_fill_gradientn(colors = PurpleAndYellow()) + 
  theme(legend.position = "right", panel.spacing=unit(0, "lines"), axis.title = element_blank(), axis.text = element_text(angle = 90, hjust = 1)) + 
  scale_size(range = c(0,20)) +
  labs(tag = "C", title = "In Vivo and In Vitro GPC Receptor Expression of Curated Ligands", size = "% Expressed", fill = "Scaled Expression") + 
    guides(colour = guide_colorbar(title.position = "top", title.theme = element_text(size = axisTitleSize)), 
         size = guide_legend(title.position = "top", title.theme = element_text(size = axisTitleSize)))

receptorGG

```

## Combined Plots
```{r}
mouse <- readRDS("output/RDS/mouse.rds")

mouseDim <- DimPlotCustom(mouse, group.by = "cellType", label = T) + theme_bw() + theme_manuscript + theme(legend.position = "bottom") + labs(tag = "A")

# This will limit which come from the next function
mouseMarkers <- c("Pdgfra", "Ptprz1", 
                  "Gpr17", "Bcas1", 
                  "Nkx6-2", "Mog",
                  "Gfap", "Aqp4",
                  "P2ry12", "Itgam",
                  "Cldn5", "Pecam1",
                  "Acta2", "Des",
                  "Dlx2", "Elavl4",
                  "Pf4", "Cd163",
                  "Tmem212", "Ccdc153")

#You can extract percent expression from this seurat function... providing features makes it way faster. $data is what you want 
mouseDotPlot <- DotPlot(mouse, features = mouseMarkers)$data

# For ordering in the plot
mouseLevels <- c("GPC", "imOL", "maOL", "Astrocyte", "Microglia", "Endothelial", "Pericyte", "NPC", "Macrophage", "Ependymal")
mouseDotPlot$id <- factor(mouseDotPlot$id , levels = rev(mouseLevels))





figMouseB <- ggplot(mouseDotPlot, aes(size = pct.exp, color = avg.exp.scaled, y = id, x = features.plot)) + 
  geom_point() + 
  scale_size_area() + 
  viridis::scale_color_viridis() + 
  theme_bw() + 
  theme_manuscript +
  theme(axis.title = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), legend.position = "bottom") +
  labs(tag = "B", title = "Canonical Marker Expression", size = "% Expressed", colour = "Scaled Expression") + 
    guides(colour = guide_colorbar(title.position = "top", title.theme = element_text(size = axisTitleSize)), 
         size = guide_legend(title.position = "top", title.theme = element_text(size = axisTitleSize))) +
  scale_size(range = c(0,4))

figMouseB

top <- (mouseDim | figMouseB) + plot_layout(widths = c(1,1.5))


middle <- (ligandActivityGG | ligandExpressionGG) +  plot_layout(widths = c(.5,10))
middle
 
(top / middle / receptorGG )

ggsave("output/Figures/Nichenet/nichenet_figure.pdf", width = 8.5, height = 12)



```

## Ligand, receptor, target Network construction
```{r}
# 
# invivoNetwork <- ligand_activities_targets[ligand_activities_targets$ligand %in% tempDF$ligand,]
# invivoNetwork <- invivoNetwork[invivoNetwork$receiver == "In Vivo GPC4",]
# 
# sigRegulons <- read.csv("output/DE/sigRegulons.csv")
# sigRegulons <- sigRegulons[sigRegulons$Gene_Log2FC >0 & sigRegulons$AUC_Log2FC > 0,]
# 
# TF_Functions <- read.csv("data_for_import/TF_Functions.csv")
# 
# sigRegulons[sigRegulons$Gene %in% TF_Functions$Repressors,]
# 
# 
# oligoTargets <- c("APOD", "S100B", "CCND1", "PTPRZ1", "MT3", "BCAS1", "OLIG1", "CNTN1", "CSPG4", "TNR", "LUZP2", "PCDH9", "BCAN", "FABP7", "CA10", "NRXN1", "PLLP", "PLP1", "BAMBI", "MBP", "OLIG2", "OLIG1", "ITM2B", "NKX2-2", "CNP")
# 
# networkTargets <- unique(c(as.character(tempDF$ligand), tempDF$receptor, 
#                            sigRegulons$Gene, oligoTargets))
# 
# invivoNetworkFilt <- invivoNetwork[invivoNetwork$target %in% networkTargets,]
# 
# invivoNodes <- data.frame(node = unique(c(invivoNetworkFilt$ligand, invivoNetworkFilt$target)))
# 
# invivoNodes$type <- "Gene Target"
# invivoNodes$type <- ifelse(invivoNodes$node %in% tempDF$receptor, "Receptor", invivoNodes$type)
# invivoNodes$type <- ifelse(invivoNodes$node %in% tempDF$ligand, "Ligand", invivoNodes$type)
# invivoNodes$type <- ifelse(invivoNodes$node %in% sigRegulons$Gene, "Regulon", invivoNodes$type)
# invivoNodes <- merge(invivoNodes, deGPC4.sig, by.x = "node", by.y = "gene", all.x = T)
# invivoNodes <- invivoNodes[,c(1,2,5)]
# 
# 
# 
# 
# 
# 
# write.table(invivoNetworkFilt, "output/Networks/NicheNet/invivoNetworkNichenet.txt", sep = "\t", quote = F, row.names = F)
# write.table(invivoNodes, "output/Networks/NicheNet/invivoNodesNichenet.txt", sep = "\t", quote = F, row.names = F)


```

## Ligand, receptor, target Network construction


```{r}

invivoNetwork <- ligand_activities_targets[ligand_activities_targets$ligand %in% tempDF$ligand,]
invivoNetwork <- invivoNetwork[invivoNetwork$receiver == "In Vivo GPC4",]

sigRegulons <- read.csv("output/DE/sigRegulons.csv")
sigRegulons <- sigRegulons[sigRegulons$Gene_Log2FC >0 & sigRegulons$AUC_Log2FC > 0,]


oligoTargets <- c("KLF6", "CCND1", "MBP", "CSPG4", "SULF2", "PCDH9", "BCAS1", "GPR17", "SEMA3E", "CA10", "OMG", "PLP1", "BAMBI", "CSPG5", "S100B", "APOD", "NRXN1", "TNR", "PLLP", "CNP", "CNTN1", "SOX10", "OLIG1", "OLIG2", "NKX2-2","PTPRZ1", "LUZP2" )

regulonEdges <- invivoNetwork[invivoNetwork$target %in% sigRegulons$Gene,]
regulonEdges$type <- "Regulon"
regulonEdges$tempLigand <- paste0(regulonEdges$ligand, "_", regulonEdges$type)
regulonEdges$tempTarget <- paste0(regulonEdges$target, "_", regulonEdges$type)
regulonNodes <- data.frame(node = unique(c(regulonEdges$ligand, regulonEdges$target)),
                           network = "Regulon")


oligoEdges <- invivoNetwork[invivoNetwork$target %in% oligoTargets,]
oligoEdges$type <- "Oligo"
oligoEdges$tempLigand <- paste0(oligoEdges$ligand, "_", oligoEdges$type)
oligoEdges$tempTarget <- paste0(oligoEdges$target, "_", oligoEdges$type)
oligoNodes <- data.frame(node = unique(c(oligoEdges$ligand, oligoEdges$target)),
                           network = "Oligo")


receptorEdges <- invivoNetwork[invivoNetwork$target %in% tempDF$receptor,]
receptorEdges$type <- "Receptor"
receptorEdges$tempLigand <- paste0(receptorEdges$ligand, "_", receptorEdges$type)
receptorEdges$tempTarget <- paste0(receptorEdges$target, "_", receptorEdges$type)
receptorNodes <- data.frame(node = unique(c(receptorEdges$ligand, receptorEdges$target)),
                           network = "Receptor")

invivoNetworkFilt <- rbindlist(list(regulonEdges, oligoEdges, receptorEdges))
invivoNetworkFilt <- invivoNetworkFilt[invivoNetworkFilt$ligand_target_weight > 0.05,]


invivoNodes <- rbindlist(list(regulonNodes, oligoNodes, receptorNodes))

invivoNodes$type <- "Gene Target"
invivoNodes$type <- ifelse(invivoNodes$node %in% tempDF$receptor, "Receptor", invivoNodes$type)
invivoNodes$type <- ifelse(invivoNodes$node %in% tempDF$ligand, "Ligand", invivoNodes$type)
invivoNodes$type <- ifelse(invivoNodes$node %in% sigRegulons$Gene, "Regulon", invivoNodes$type)
invivoNodes <- merge(invivoNodes, deGPC4.sig, by.x = "node", by.y = "gene", all.x = T)
invivoNodes <- invivoNodes[,c(1,2,3,6)]
invivoNodes$tempNode <- paste0(invivoNodes$node, "_", invivoNodes$network)
invivoNodes

TF_Functions <- read.csv("data_for_import/TF_Functions.csv")

invivoNodes[invivoNodes$type == "Regulon" & invivoNodes$node %not in% c(TF_Functions$Activators, TF_Functions$Repressors),]

invivoNodes <- invivoNodes[invivoNodes$node %not in% TF_Functions$Repressors,]

invivoNetworkFilt <- invivoNetworkFilt[invivoNetworkFilt$target %not in% TF_Functions$Repressors,]


write.table(invivoNetworkFilt, "output/Networks/NicheNet/invivoNetworkNichenet.txt", sep = "\t", quote = F, row.names = F)
write.table(invivoNodes, "output/Networks/NicheNet/invivoNodesNichenet.txt", sep = "\t", quote = F, row.names = F)



```

# Supplementary
```{r}



dimLeidenFig <- DimPlotCustom(mouse, group.by = "leidenClusters", ncol = 1, label = T, pt.size = 3) & theme_bw() & theme_manuscript & NoLegend()  & ggtitle("In Vivo Mouse Leiden Clusters") & labs(tag = "A")

dimLeidenFig

```


## Leiden Markers
```{r}

DefaultAssay(mouse) <- "RNA"

canonicalMarkers <- c("Pdgfra", "Ptprz1", 
                  "Gpr17", "Bcas1", 
                  "Nkx6-2", "Mog",
                  "Gfap", "Aqp4",
                  "P2ry12", "Itgam",
                  "Cldn5", "Pecam1",
                  "Acta2", "Des",
                  "Dlx2", "Elavl4",
                  "Pf4", "Cd163",
                  "Tmem212", "Ccdc153")

markerDotPlotSupp <- DotPlot(mouse, features = canonicalMarkers, group.by = "leidenClusters")$data

leidenTemp <- c("0" = "maOL",
                          "1" = "maOL",
                          "2" = "maOL",
                          "3" = "maOL",
                          "4" = "maOL",
                          "5" = "maOL",
                          "6" = "maOL",
                          "7" = "Endothelial",
                          "8" = "maOL",
                          "9" = "maOL",
                          "10" = "maOL",
                          "11" = "maOL",
                          "12" = "maOL",
                          "13" = "Microglia",
                          "14" = "imOL",
                          "15" = "GPC",
                          "16" = "Astrocyte",
                          "17" = "NPC",
                          "18" = "Ependymal",
                          "19" = "Pericyte",
                          "20" = "Macrophage")


leidenTemp <- paste0(0:20, " - ", leidenTemp)
names(leidenTemp) <- 0:20

markerDotPlotSupp$id <- plyr::mapvalues(as.character(markerDotPlotSupp$id), from = names(leidenTemp), to = leidenTemp)

markerDotPlotSupp$id <- factor(markerDotPlotSupp$id, levels = rev(c("15 - GPC",
                                                                    "14 - imOL",
                                                                    "5 - maOL",
                                                                    "0 - maOL",
                                                                    "1 - maOL",
                                                                    "2 - maOL",
                                                                    "3 - maOL",
                                                                    "4 - maOL",
                                                                    "6 - maOL",
                                                                    "8 - maOL",
                                                                    "9 - maOL",
                                                                    "10 - maOL",
                                                                    "11 - maOL",
                                                                    "12 - maOL",
                                                                    "16 - Astrocyte",
                                                                    "13 - Microglia",
                                                                    "7 - Endothelial",
                                                                    "19 - Pericyte",
                                                                    "17 - NPC",
                                                                    "20 - Macrophage",
                                                                    "18 - Ependymal")))



figSuppMarkerPlot<- ggplot(markerDotPlotSupp, aes(size = pct.exp, fill = avg.exp.scaled, y = id, x = features.plot)) +
  geom_point(color = "black", pch = 21) + 
  scale_size_area(max_size = 15) + 
  scale_fill_gradientn(colors = PurpleAndYellow()) + 
  theme_bw() + 
  theme_manuscript +
  theme(axis.title = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), legend.position = "right", panel.spacing=unit(0, "lines")) +
  labs(tag = "B", title = "Canonical Marker Expression", size = "% Expressed", fill = "Scaled Expression") + 
    guides(colour = guide_colorbar(title.position = "top", title.theme = element_text(size = axisTitleSize)), 
         size = guide_legend(title.position = "top", title.theme = element_text(size = axisTitleSize)))


figSuppMarkerPlot


```


## Piece together

```{r}

((dimLeidenFig | figSuppMarkerPlot) / receptorGG) + plot_layout(heights = c(1,.25))

ggsave("output/Figures/Nichenet/nichenetSupplement.pdf", width = 30, height = 20)


```


```{r}

sessionInfo()

```

